---
title: "Appendix C.II: Split-Sample Method - Confirmatory Latent Class Analysis (CLCA)"
author: "Adam Garber"
date: 'Updated: October 2, 2024'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE, tidy = TRUE)
```

```{r}
library(tidyverse)
library(MplusAutomation)
library(rhdf5)
library(here)
library(glue)
library(gt)
library(janitor)
library(reshape2)
library(cowplot)
```

Read data file `n_6000_lca_rep1.dat` (N=6000; Replication 1)
```{r}
lca_data <- read.delim2(here("C1-Simulation", "n_6000_lca_rep1.dat"),
  sep = "", header = FALSE) %>%
  select(-V18) %>% 
  setNames(
    c("primary", "change", "interupt", "initiat", "engage", "approach", "response", "expect",
      "new", "same", "relative", "objects", "sequence", "trans", "avoid", "control", "touch")) %>% 
  purrr::modify_if(is.character, as.numeric) 
  
```

```{r, eval=FALSE}

write_csv(lca_data, here("data", "C2_simulated_N6000R1.csv"))

```

Randomly select rows to split the parent sample equally into exploratory and confirmatory sub-samples

- Exploratory sample (`explr_data`): N = 3,000
- Confirmatory sample (`cnfrm_data`): N = 3,000

```{r}
set.seed(9182024) # For reproducibility

split_prop <- 0.5

split_data <- lca_data %>%
  mutate(split = ifelse(row_number() %in% sample(1:n(), size = floor(split_prop * n())), 
                        "exploratory", "confirmatory"))

# Create sub-samples
explr_data <- split_data %>% filter(split == "exploratory") %>% select(-split)
cnfrm_data  <- split_data %>% filter(split == "confirmatory") %>% select(-split)

```

------------------------------------------------------------------------

Step 1

------------------------------------------------------------------------

Estimate exploratory LCA model with data subset `explr_data` (N=3000)
```{r, eval=FALSE}

exploratoryLCA  <- lapply(4:6, function(k) {
  lca_enum  <- mplusObject(
      
    TITLE = glue("{k}-Class"), 
  
    VARIABLE = glue(
    "!!! Split Sample Method (STEP-1): Exploratory Model !!!
     categorical = primary-touch; 
     usevar = primary-touch; 
     classes = c({k}); "),
  
  ANALYSIS = 
   "estimator = mlr; 
    type = mixture;
    starts = 500 200; 
    processors = 10;",
  
  OUTPUT = "svalues residual tech11 tech14;",
  
  PLOT = 
    "type = plot3; 
    series = primary-touch(*);",
  
  usevariables = colnames(explr_data),
  rdata = explr_data)

lca_enum_fit <- mplusModeler(lca_enum, 
                     dataout=glue(here("C2-Split-Sample", "lca_explr_data.dat")),
                     modelout=glue(here("C2-Split-Sample", "c{k}_explr_data.inp")),
                     check=TRUE, run = TRUE, hashfilename = FALSE)
})

```

------------------------------------------------------------------------

### Table of Fit

Extract fit data
```{r}

output_lca <- readModels(here("C2-Split-Sample"), filefilter = "explr", quiet = TRUE)

enum_extract <- LatexSummaryTable(output_lca,
    keepCols = c("Title","Parameters","LL","BIC","aBIC",
    "BLRT_PValue","T11_VLMR_PValue","Observations"),
    sortBy = "Title") 

allFit <- enum_extract %>%
  mutate(CAIC = -2 * LL + Parameters * (log(Observations) + 1)) %>%
  mutate(AWE = -2 * LL + 2 * Parameters * (log(Observations) + 1.5)) %>%
  mutate(SIC = -.5 * BIC) %>%
  mutate(expSIC = exp(SIC - max(SIC))) %>%
  mutate(BF = exp(SIC - lead(SIC))) %>%
  mutate(cmPk = expSIC / sum(expSIC)) %>%
  dplyr::select(1:5, 9:10, 6:7, 13, 14) %>%
  arrange(Parameters)
```

Create table
```{r}
fit_table <- allFit %>%
  gt() %>%
  tab_header(title = md("**Model Fit Summary Table**")) %>%
  cols_label(
    Title = "Classes",
    Parameters = md("Par"),
    LL = md("*LL*"),
    T11_VLMR_PValue = "VLMR",
    BLRT_PValue = "BLRT",
    BF = md("BF"),
    cmPk = md("*cmPk*")
  ) %>%
  tab_footnote(
    footnote = md(
      "*Note.* Par = Parameters; *LL* = model log likelihood;
BIC = Bayesian information criterion;
aBIC = sample size adjusted BIC; CAIC = consistent Akaike information criterion;
AWE = approximate weight of evidence criterion;
BLRT = bootstrapped likelihood ratio test p-value;
VLMR = Vuong-Lo-Mendell-Rubin adjusted likelihood ratio test p-value;
*cmPk* = approximate correct model probability."
    ),
locations = cells_title()
  ) %>%
  tab_options(column_labels.font.weight = "bold") %>%
  fmt_number(c(3:7),
             decimals = 2) %>%
  sub_missing(1:11,
              missing_text = "--") %>%
  fmt(c(8:9, 11),
  fns = function(x)
  ifelse(x < 0.001, "<.001", scales::number(x, accuracy = .01))
  ) %>%
  fmt(10,fns = function (x)
      ifelse(x > 100, ">100", scales::number(x, accuracy = .01))
  ) %>%  
  tab_style(
    style = list(
      cell_text(weight = "bold")),
    locations = list(cells_body(
     columns = BIC,
     row = BIC == min(BIC[c(1:3)]) # Change this to the number of classes you are evaluating
    ),
    cells_body(
     columns = aBIC,
     row = aBIC == min(aBIC[1:3])
    ),
    cells_body(
     columns = CAIC,
     row = CAIC == min(CAIC[1:3])
    ),
    cells_body(
     columns = AWE,
     row = AWE == min(AWE[1:3])
    ),
    cells_body(
     columns = cmPk,
     row =  cmPk == max(cmPk[1:3])
     ),    
    cells_body(
     columns = BF,
     row =  BF > 10),
    cells_body( 
     columns =  T11_VLMR_PValue,
     row =  ifelse(T11_VLMR_PValue < .05 & lead(T11_VLMR_PValue) > .05, T11_VLMR_PValue < .05, NA)),
    cells_body(
     columns =  BLRT_PValue,
     row =  ifelse(BLRT_PValue < .05 & lead(BLRT_PValue) > .05, BLRT_PValue < .05, NA))
  )
)

fit_table
```

Save table
```{r, eval = FALSE}
gtsave(fit_table, here("figures", "C2-fit-table-explr-data.png"))

```

------------------------------------------------------------------------

### Conditional LCA Plot Function

```{r, eval=TRUE}

plot_lca_function <- function(model_name,item_num,class_num,item_labels,
                              class_labels=c("Class1","Class2","Class3","Class4", "Class5"),
                              class_legend_order=c(1,2,3,4,5),plot_title){

mplus_model <- as.data.frame(model_name$gh5$means_and_variances_data$estimated_probs$values)
plot_data <- mplus_model[seq(2, 2*item_num, 2),]

c_size <- as.data.frame(model_name$class_counts$modelEstimated$proportion)
colnames(c_size) <- paste0("cs")
c_size <- c_size %>% mutate(cs = round(cs*100, 2))
colnames(plot_data) <- paste0(class_labels, glue(" ({c_size[1:class_num,]}%)"))
plot_data <- plot_data %>% relocate(class_legend_order)

plot_data <- cbind(Var = paste0("U", 1:item_num), plot_data)
plot_data$Var <- fct_inorder(plot_data$Var)
plot_data$Var <- factor(plot_data$Var, labels = item_labels)


pd_long_data <- melt(plot_data, id.vars = "Var") 

# This syntax uses the data.frame created above to produce the plot with `ggplot()`

p <- pd_long_data %>%
  ggplot(aes(x = as.integer(Var), y = value,
  shape = variable, colour = variable, lty = variable)) +
  geom_point(size = 4) + geom_line( size=2 ) +
  scale_x_continuous("", breaks = 1:item_num, labels = plot_data$Var) + 
  scale_color_viridis_d(end=.7, alpha=.89, option = "H") +
  labs(title = plot_title, y = "Probability") +
  theme_cowplot() +
  theme(legend.title = element_blank(), 
        legend.position = "top",
        axis.text.x = element_text(size = 8, vjust = 1))
p
return(p)
}

```


------------------------------------------------------------------------

### 5-Class Probability Plot

Use the `plot_lca` function provided in the folder to plot the item probability plot. Details for this function are written in the document `plot_lca.txt`

Read models for plotting (5-class model)

```{r, eval=TRUE}
model_c5 <- readModels(here("C2-Split-Sample", "c5_explr_data.out"), quiet = TRUE)
```

```{r fig.height=6, fig.width=10}

plot_lca_function(
  model_name = model_c5, 
  item_num = 17,
  class_num = 5,
  item_labels = c("Primary","Change", "Interupt","Initiate","Engage","Approach","Response","Expect",
                "New","Same","Relative","Objects","Sequence", "Transition","Avoid","Control", "Touch"),
  class_labels = c("High flex.","Place/sens. inflex.","Env. flex.","Low flex.","Soc./intr. flex."),
  class_legend_order = c(1,5,3,2,4),
  plot_title = "Conditional Item Probability Plot"
  )

```

Save figure
```{r, eval = FALSE}
ggsave(here("figures", "C5_Plot_Explr-Data.png"), dpi=300, height=5, width=11, units="in")
```

------------------------------------------------------------------------

Step 2: Confirmatory analysis

------------------------------------------------------------------------

CLCA model estimated from the split sample `cnfrm_data` using start values from exploratory K=5 model 
```{r, eval = FALSE}

m_step1  <- mplusObject(
  
  TITLE = "Split-Sample (confirmatory sample)", 
  
  VARIABLE = 
   "categorical = primary-touch; 
    usevar = primary-touch;
    classes = c(5); ",
  
  ANALYSIS = 
   "estimator = mlr; 
    type = mixture; 
    !starts = 0;
    ! STSEED = 392407; !!! USE SEED TO REPLICATE THESIS RESULTS !!!
   ",
  
  MODEL = 
    "%OVERALL%

     [ c#1*-1.74441 ];  
     [ c#2*-0.64595 ];  
     [ c#3*-0.29794 ];  
     [ c#4*-1.17904 ];  

     %C#1%       !!!High_Flex!!!

     [ primary$1*-2.20987 ](t1);
     [ change$1*-1.84601  ](t2);
     [ interupt$1*-1.88724](t3);
     [ initiat$1*-2.13724 ](t4);
     [ engage$1*-1.86574  ](t5);
     [ approach$1*-2.38052](t6);
     [ response$1*-2.36713](t7);
     [ expect$1*-2.45601  ](t8);
     [ new$1*-3.60891     ](t9);
     [ same$1*-2.47183    ](t10);
     [ relative$1*-1.67673](t11);
     [ objects$1*-2.75974 ](t12);
     [ sequence$1*-2.08144](t13);
     [ trans$1*-2.97972   ](t14);
     [ avoid$1*-2.32553   ](t15);
     [ control$1*-1.99748 ](t16);
     [ touch$1*-1.80536   ](t17);

     %C#2%         !!!LocSnsLo!!!

     [ primary$1*-0.29794 ](t52);     
     [ change$1*-0.53213  ](t53);    
     [ interupt$1*-0.54672](t54);      
     [ initiat$1*-0.45183 ](t55);     
     [ engage$1*-0.51045  ](t56); 
     [ approach$1*-0.47146](t57);      
     [ response$1*-0.63201](t58);      
     [ expect$1*-0.67809  ](t59);    
     [ new$1*1.20090      ](t60);
     [ same$1*2.84868     ](t61); 
     [ relative$1*1.87798 ](t62);     
     [ objects$1*-1.88945 ](t63);    
     [ sequence$1*-1.45427](t64);     
     [ trans$1*-1.29703   ](t65);  
     [ avoid$1*1.03081    ](t66);  
     [ control$1*1.63541  ](t67);    
     [ touch$1*1.29558    ](t68);  

     %C#3%       !!!Env_Flex !!!

     [ primary$1*3.76057  ](t35);      
     [ change$1*1.23660   ](t36);     
     [ interupt$1*2.42362 ](t37);       
     [ initiat$1*1.07970  ](t38);      
     [ engage$1*4.47606   ](t39);     
     [ approach$1*1.51131 ](t40);       
     [ response$1*2.02784 ](t41);       
     [ expect$1*2.53496   ](t42);     
     [ new$1*-2.44811     ](t43);  
     [ same$1*-0.99306    ](t44);   
     [ relative$1*-1.67001](t45);       
     [ objects$1*-0.95871 ](t46); 
     [ sequence$1*-0.88720](t47);       
     [ trans$1*-0.92516   ](t48);    
     [ avoid$1*-1.17745   ](t49);    
     [ control$1*-1.78766 ](t50);      
     [ touch$1*-0.99513   ](t51);  

     %C#4%         !!!Low_Flex!!! 

     [ primary$1*3.37636  ](t69);      
     [ change$1*1.81614   ](t70);     
     [ interupt$1*2.02582 ](t71);       
     [ initiat$1*3.62826  ](t72);      
     [ engage$1*2.72187   ](t73);     
     [ approach$1*3.20886 ](t74);       
     [ response$1*2.96584 ](t75);       
     [ expect$1*2.92061   ](t76);     
     [ new$1*2.06398      ](t77);
     [ same$1*2.88648     ](t78);  
     [ relative$1*2.69372 ](t79);      
     [ objects$1*3.14932  ](t80);      
     [ sequence$1*2.13968 ](t81);       
     [ trans$1*1.75316    ](t82);    
     [ avoid$1*1.73877    ](t83);   
     [ control$1*2.78570  ](t84);     
     [ touch$1*2.99815    ](t85);   

     %C#5%       !!!ScIn_Flex!!! 

     [ primary$1*-1.11218 ](t18);     
     [ change$1*-0.98964  ](t19);    
     [ interupt$1*-1.53552](t20);      
     [ initiat$1*-1.91915 ](t21);     
     [ engage$1*-0.92521  ](t22);    
     [ approach$1*-1.33903](t23);      
     [ response$1*-1.53825](t24);      
     [ expect$1*-1.76776  ](t25);    
     [ new$1*1.23723      ](t26); 
     [ same$1*1.32099     ](t27);  
     [ relative$1*0.95454 ](t28);      
     [ objects$1*2.16596  ](t29);     
     [ sequence$1*2.55756 ](t30);      
     [ trans$1*1.84039    ](t31);   
     [ avoid$1*1.24333    ](t32);   
     [ control$1*3.37130  ](t33);     
     [ touch$1*2.06513    ](t34);   
  ",
  
  MODELCONSTRAINT = "!!!       THRESHOLD BOUNDARIES            !!!
                     !!! -.85 (THRESHOLD) ~ .70 (PROBABILTY)   !!!
                     !!!  .85 (THRESHOLD) ~ .30 (PROBABILTY)   !!!
      
     ! LABELS C1-C5 BELOW REFLECT ORDER OF CLASSES IN PLOT (NOT MPLUS C# LABELS ABOVE)
     !    C1   |    C2   |   C3    |    C4   |   C5    !
     !High_Flex|ScIn_Flex|Env_Flex |LocSnsLo |Low_Flex !
     !_________|_________|_________|_________|_________!
      t1 <-.85; t18<-.85; t35> .85;           t69> .85;  !!!U1 !!!
      t2 <-.85; t19<-.85; t36> .85;           t70> .85;  !!!U2 !!!
      t3 <-.85; t20<-.85; t37> .85;           t71> .85;  !!!U3 !!!
      t4 <-.85; t21<-.85; t38> .85;           t72> .85;  !!!U4 !!!
      t5 <-.85; t22<-.85; t39> .85;           t73> .85;  !!!U5 !!!
      t6 <-.85; t23<-.85; t40> .85;           t74> .85;  !!!U6 !!!
      t7 <-.85; t24<-.85; t41> .85;           t75> .85;  !!!U7 !!!
      t8 <-.85; t25<-.85; t42> .85;           t76> .85;  !!!U8 !!!
      
      t9 <-.85; t26> .85; t43<-.85; t60> .85; t77> .85;  !!!U9 !!!
      t10<-.85; t27> .85; t44<-.85; t61> .85; t78> .85;  !!!U10!!!
      t11<-.85; t28> .85; t45<-.85; t62> .85; t79> .85;  !!!U11!!!
      t12<-.85; t29> .85; t46<-.85; t63<-.85; t80> .85;  !!!U12!!!
      t13<-.85; t30> .85; t47<-.85; t64<-.85; t81> .85;  !!!U13!!!
      t14<-.85; t31> .85; t48<-.85; t65<-.85; t82> .85;  !!!U14!!!
      t15<-.85; t32> .85; t49<-.85; t66> .85; t83> .85;  !!!U15!!!
      t16<-.85; t33> .85; t50<-.85; t67> .85; t84> .85;  !!!U16!!!
      t17<-.85; t34> .85; t51<-.85; t68> .85; t85> .85;  !!!U17!!!
  
       !!! EQUALITY (& INEQUALITY) CONSTRAINTS !!!
       !  C1vC2  |  C3vC4  |  C4vC5  !
       !_________|_________|_________!
        t1 = t18; t35=-t52;           !!!U1 !!!
        t2 = t19; t36=-t53;           !!!U2 !!!
        t3 = t20; t37=-t54;           !!!U3 !!!
        t4 = t21; t38=-t55;           !!!U4 !!!
        t5 = t22; t39=-t56;           !!!U5 !!!
        t6 = t23; t40=-t57;           !!!U6 !!!
        t7 = t24; t41=-t58;           !!!U7 !!!
        t8 = t25; t42=-t59;           !!!U8 !!!
        t9 =-t26;                     !!!U9 !!!
        t10=-t27;                     !!!U10!!!
        t11=-t28;                     !!!U11!!!
        t12=-t29;           t63=-t80; !!!U12!!!
        t13=-t30;           t64=-t81; !!!U13!!!
        t14=-t31;           t65=-t82; !!!U14!!!
        t15=-t32;                     !!!U15!!!
        t16=-t33;                     !!!U16!!!
        t17=-t34;                     !!!U17!!!
       
!--------------------------------------------------------------!

  ",
  
  OUTPUT = "tech14;",

  PLOT = 
    "type = plot3; 
    series = primary-touch(*);",
  
  usevariables = colnames(cnfrm_data),
  rdata = cnfrm_data)

m_step1_fit <- mplusModeler(m_step1,
                 dataout=here("C2-Split-Sample", "CLCA_SplitSample.dat"),
                 modelout=here("C2-Split-Sample", "CLCA_SplitSample.inp") ,
                 check=TRUE, run = TRUE, hashfilename = FALSE)

```

------------------------------------------------------------------------

Step 3: Compare model fit

------------------------------------------------------------------------

Conduct the Sattorra-Bentler adjusted Log Likelihood Ratio (LRT) difference test:

- Exploratory model (parent): `c5_explr_data.out` is the 'un-constrained model" with 89 parameters.

- Confirmatory model (nested): `CLCA_SplitSample.out` is the "constrained model" with 61 parameters.

```{r}
split_sample_models <- readModels(here("C2-Split-Sample"), quiet = TRUE)
```

```{r}
# *0 = null or nested model & *1 = comparison or parent model

# Log Likelihood Values
L0 <- split_sample_models[["CLCA_SplitSample.out"]][["summaries"]][["LL"]]
L1 <- split_sample_models[["c5_explr_data.out"]][["summaries"]][["LL"]] 

# LRT equation
lr <- -2*(L0-L1) 

# Parameters
p0 <- split_sample_models[["CLCA_SplitSample.out"]][["summaries"]][["Parameters"]] 
p1 <- split_sample_models[["c5_explr_data.out"]][["summaries"]][["Parameters"]]

# Scaling Correction Factors
c0 <- split_sample_models[["CLCA_SplitSample.out"]][["summaries"]][["LLCorrectionFactor"]]
c1 <- split_sample_models[["c5_explr_data.out"]][["summaries"]][["LLCorrectionFactor"]]

# Difference Test Scaling correction (Sattorra-Bentler adjustment)
cd <- ((p0*c0)-(p1*c1))/(p0-p1)

# Chi-square difference test(TRd)
TRd <- (lr)/(cd)

# Degrees of freedom
df <- abs(p0 - p1)


# Significance test
(p_diff <- pchisq(TRd, df, lower.tail=FALSE))

```

**RESULT**: The Log Likelihood $\chi^2$ difference test comparing the exploratory and confirmatory LCA models was, $\chi^2 (28) = 1588.06, p < .001$. [\textcolor{blue}{See Reference Here}](http://www.statmodel.com/chidiff.shtml)

------------------------------------------------------------------------

Compare model fit summary statistics: Exploratory & Confirmatory Models 
```{r, eval=TRUE}

enum_extract1 <- LatexSummaryTable(split_sample_models$c5_explr_data.out ,                                 
                 keepCols=c("Title", "Parameters", "LL", "BIC", "aBIC", "Observations")) 
enum_extract2 <- LatexSummaryTable(split_sample_models$CLCA_SplitSample.out,                                 
                 keepCols=c("Title", "Parameters", "LL", "BIC", "aBIC", "Observations")) 
```

Calculate indices derived from the Log Likelihood (LL)
```{r, eval=TRUE}
                           
allFit <- rbind(enum_extract1, enum_extract2) %>% 
  mutate(aBIC = -2*LL+Parameters*log((Observations+2)/24)) %>% 
  select(1:5) %>% 
  mutate(Title = case_when(
    Title == " 5-Class" ~ "Exploratory Model",
    Title == " Split-Sample (confirmatory sample)" ~ "Confirmatory Model"))

```

Format fit table
```{r, eval=TRUE}

allFit %>% 
  gt() %>%
  tab_header(
    title = md("**Model Fit Comparision Table**"), subtitle = md("&nbsp;")) %>% 
  cols_label(
    Title = "Model",
    Parameters = md("Par"),
    LL = md("*LL*"),
    BIC = md("BIC"),
    aBIC = md("aBIC")) %>% 
  tab_footnote(
    footnote = md(
    "*Note.* Par = Parameters; *LL* = model log likelihood;
     BIC = Bayesian information criterion; aBIC = sample size adjusted BIC."), 
    locations = cells_title()) %>% 
  tab_options(column_labels.font.weight = "bold")
```
















