---
title: "Appendix C.IV: Replication Method - Confirmatory Latent Class Analysis (CLCA)"
author: "Adam Garber"
date: 'Updated: October 2, 2024'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE, tidy = TRUE)
```

```{r}
library(tidyverse)
library(MplusAutomation)
library(rhdf5)
library(here)
library(glue)
library(gt)
library(janitor)
library(reshape2)
library(cowplot)
```

Read in data files for the initial sample and replication sample:

- Initial sample: `n_3000_lca_rep2.dat` (N=3000; Replication 2)
- Replication sample: `n_3000_lca_rep3.dat` (N=3000; Replication 3)
```{r}
# Read in initial sample
sample_1 <- read.delim2(here("C1-Simulation", "n_3000_lca_rep2.dat"),
  sep = "", header = FALSE) %>%
  select(-V18) %>% 
  setNames(
    c("primary", "change", "interupt", "initiat", "engage", "approach", "response", "expect",
      "new", "same", "relative", "objects", "sequence", "trans", "avoid", "control", "touch")) %>% 
  purrr::modify_if(is.character, as.numeric) %>% 
  mutate(sample = 1)

sample_2 <- read.delim2(here("C1-Simulation", "n_3000_lca_rep3.dat"),
  sep = "", header = FALSE) %>%
  select(-V18) %>% 
  setNames(
    c("primary", "change", "interupt", "initiat", "engage", "approach", "response", "expect",
      "new", "same", "relative", "objects", "sequence", "trans", "avoid", "control", "touch")) %>% 
  purrr::modify_if(is.character, as.numeric) %>% 
  mutate(sample = 2)
```

Combine samples 1 and 2 for multigroup analysis
```{r}

combined_data <- rbind(sample_1, sample_2)

```

Save the `combined_data` 
```{r, eval=FALSE}

write_csv(combined_data, here("data", "C4_Combined_Data.csv"))

```

------------------------------------------------------------------------

Step 1

------------------------------------------------------------------------

Produce start values from the initial sample (`sample_1`) to use in Step 2
```{r, eval = FALSE}

m_start  <- mplusObject(
  
  TITLE = "Starts - Sample 1", 
  
  VARIABLE = 
   "categorical = primary-touch; 
    usevar = primary-touch;
    classes = c(5); ",
  
  ANALYSIS = 
   "estimator = mlr; 
    type = mixture; 
    starts = 500 200;
    !STSEED = 887676; !!! USE SEED TO REPLICATE THESIS RESULTS !!!
   ",
  
  OUTPUT = "svalues;",

  PLOT = "type = plot3; 
          series = primary-touch(*);",
  
  usevariables = colnames(sample_1),
  rdata = sample_1)

m_start_fit <- mplusModeler(m_start,
                 dataout=here("C4-Replication", "Starts_Sample_1.dat"),
                 modelout=here("C4-Replication", "Starts_Sample_1.inp") ,
                 check=TRUE, run = TRUE, hashfilename = FALSE)

```

------------------------------------------------------------------------

Step 2

------------------------------------------------------------------------

Estimate baseline multigroup model with thresholds unconstrained across groups
```{r, eval = FALSE}

m_step2  <- mplusObject(
  
  TITLE = "Step 2 - Replication", 
  
  VARIABLE = 
   "categorical = primary-touch; 
    usevar = primary-touch;
  
    knownclass = CG (sample=1 sample=2); !!! LABEL SAMPLES !!!
    classes = CG(2) C(5);                !!! CG is KNOWN CLASS !!!
   ",
  
  ANALYSIS = 
   "estimator = mlr; 
    type = mixture;
    starts = 500 200;
    !STSEED = ; !!! USE SEED TO REPLICATE THESIS RESULTS !!!
   ",
  
  MODEL = 
    "%OVERALL%
     
     !!! INITIAL SAMPLE (SAMPLE 1) !!! INITIAL SAMPLE (SAMPLE 1) !!!
  
     %CG#1.C#1% 

     [ primary$1*-0.21760  ];             
     [ change$1*-0.45914   ];            
     [ interupt$1*-0.64034 ];              
     [ initiat$1*-0.46809  ];             
     [ engage$1*-0.33510   ];            
     [ approach$1*-0.42163 ];              
     [ response$1*-0.58459 ];              
     [ expect$1*-0.60368   ];            
     [ new$1*1.17714       ];        
     [ same$1*2.92160      ];         
     [ relative$1*1.86782  ];             
     [ objects$1*-1.97843  ];             
     [ sequence$1*-1.40431 ];              
     [ trans$1*-1.31741    ];           
     [ avoid$1*0.99669     ];          
     [ control$1*1.69313   ];            
     [ touch$1*1.24603     ];          

     %CG#1.C#2%

     [ primary$1*3.45112   ];        
     [ change$1*2.10148    ];       
     [ interupt$1*2.33173  ];         
     [ initiat$1*4.67053   ];        
     [ engage$1*2.61317    ];       
     [ approach$1*3.00689  ];         
     [ response$1*3.01867  ];         
     [ expect$1*2.47608    ];       
     [ new$1*2.13464       ];    
     [ same$1*2.61928      ];     
     [ relative$1*2.46710  ];         
     [ objects$1*3.49010   ];        
     [ sequence$1*2.16238  ];         
     [ trans$1*1.76331     ];      
     [ avoid$1*1.62916     ];      
     [ control$1*2.77768   ];        
     [ touch$1*2.98188     ];      

     %CG#1.C#3%

     [ primary$1*-1.15805  ];       
     [ change$1*-1.05640   ];      
     [ interupt$1*-1.47052 ];        
     [ initiat$1*-1.88494  ];       
     [ engage$1*-0.85925   ];      
     [ approach$1*-1.29676 ];        
     [ response$1*-1.60551 ];        
     [ expect$1*-1.84879   ];      
     [ new$1*1.21384       ];  
     [ same$1*1.25810      ];   
     [ relative$1*1.00299  ];       
     [ objects$1*2.02420   ];      
     [ sequence$1*2.49353  ];       
     [ trans$1*1.91476     ];    
     [ avoid$1*1.22229     ];    
     [ control$1*3.67528   ];      
     [ touch$1*2.02070     ];    

     %CG#1.C#4%

     [ primary$1*-2.45913  ];           
     [ change$1*-2.07089   ];          
     [ interupt$1*-2.29483 ];            
     [ initiat$1*-2.28241  ];           
     [ engage$1*-1.87192   ];          
     [ approach$1*-2.25327 ];            
     [ response$1*-2.27708 ];            
     [ expect$1*-2.29726   ];          
     [ new$1*-2.86563      ];       
     [ same$1*-2.02751     ];        
     [ relative$1*-2.09303 ];            
     [ objects$1*-3.25774  ];           
     [ sequence$1*-2.18406 ];            
     [ trans$1*-2.90318    ];         
     [ avoid$1*-2.16937    ];         
     [ control$1*-1.95077  ];           
     [ touch$1*-1.64226    ];         

     %CG#1.C#5%

     [ primary$1*3.92177   ];     
     [ change$1*1.18687    ];    
     [ interupt$1*2.44270  ];      
     [ initiat$1*1.17768   ];     
     [ engage$1*4.46855    ];    
     [ approach$1*1.51338  ];      
     [ response$1*2.11385  ];      
     [ expect$1*2.56064    ];    
     [ new$1*-2.43669      ];  
     [ same$1*-1.04154     ];   
     [ relative$1*-1.78324 ];       
     [ objects$1*-1.14240  ];      
     [ sequence$1*-0.84974 ];       
     [ trans$1*-0.89213    ];    
     [ avoid$1*-1.11250    ];    
     [ control$1*-1.78770  ];      
     [ touch$1*-0.87239    ]; 
  
     !!! REPLICATION SAMPLE (SAMPLE 2) !!! REPLICATION SAMPLE (SAMPLE 2) !!!
  
     %CG#2.C#1% 

     [ primary$1*-0.21760  ];            
     [ change$1*-0.45914   ];           
     [ interupt$1*-0.64034 ];             
     [ initiat$1*-0.46809  ];            
     [ engage$1*-0.33510   ];           
     [ approach$1*-0.42163 ];             
     [ response$1*-0.58459 ];             
     [ expect$1*-0.60368   ];           
     [ new$1*1.17714       ];       
     [ same$1*2.92160      ];         
     [ relative$1*1.86782  ];             
     [ objects$1*-1.97843  ];             
     [ sequence$1*-1.40431 ];              
     [ trans$1*-1.31741    ];           
     [ avoid$1*0.99669     ];          
     [ control$1*1.69313   ];            
     [ touch$1*1.24603     ];          

     %CG#2.C#2%

     [ primary$1*3.45112   ];        
     [ change$1*2.10148    ];       
     [ interupt$1*2.33173  ];         
     [ initiat$1*4.67053   ];        
     [ engage$1*2.61317    ];       
     [ approach$1*3.00689  ];         
     [ response$1*3.01867  ];         
     [ expect$1*2.47608    ];       
     [ new$1*2.13464       ];    
     [ same$1*2.61928      ];     
     [ relative$1*2.46710  ];         
     [ objects$1*3.49010   ];        
     [ sequence$1*2.16238  ];         
     [ trans$1*1.76331     ];      
     [ avoid$1*1.62916     ];      
     [ control$1*2.77768   ];        
     [ touch$1*2.98188     ];      

     %CG#2.C#3%

     [ primary$1*-1.15805  ];       
     [ change$1*-1.05640   ];      
     [ interupt$1*-1.47052 ];        
     [ initiat$1*-1.88494  ];       
     [ engage$1*-0.85925   ];      
     [ approach$1*-1.29676 ];        
     [ response$1*-1.60551 ];        
     [ expect$1*-1.84879   ];      
     [ new$1*1.21384       ];  
     [ same$1*1.25810      ];   
     [ relative$1*1.00299  ];       
     [ objects$1*2.02420   ];      
     [ sequence$1*2.49353  ];       
     [ trans$1*1.91476     ];    
     [ avoid$1*1.22229     ];    
     [ control$1*3.67528   ];      
     [ touch$1*2.02070     ];    

     %CG#2.C#4%

     [ primary$1*-2.45913  ];       
     [ change$1*-2.07089   ];      
     [ interupt$1*-2.29483 ];        
     [ initiat$1*-2.28241  ];       
     [ engage$1*-1.87192   ];      
     [ approach$1*-2.25327 ];        
     [ response$1*-2.27708 ];        
     [ expect$1*-2.29726   ];      
     [ new$1*-2.86563      ];   
     [ same$1*-2.02751     ];    
     [ relative$1*-2.09303 ];        
     [ objects$1*-3.25774  ];       
     [ sequence$1*-2.18406 ];        
     [ trans$1*-2.90318    ];     
     [ avoid$1*-2.16937    ];     
     [ control$1*-1.95077  ];       
     [ touch$1*-1.64226    ];     

     %CG#2.C#5%

     [ primary$1*3.92177   ];     
     [ change$1*1.18687    ];    
     [ interupt$1*2.44270  ];      
     [ initiat$1*1.17768   ];     
     [ engage$1*4.46855    ];    
     [ approach$1*1.51338  ];      
     [ response$1*2.11385  ];      
     [ expect$1*2.56064    ];    
     [ new$1*-2.43669      ];  
     [ same$1*-1.04154     ];   
     [ relative$1*-1.78324 ];       
     [ objects$1*-1.14240  ];      
     [ sequence$1*-0.84974 ];       
     [ trans$1*-0.89213    ];    
     [ avoid$1*-1.11250    ];    
     [ control$1*-1.78770  ];      
     [ touch$1*-0.87239    ];   ",
  
  
  OUTPUT = "",

  PLOT = "type = plot3; 
          series = primary-touch(*);",
  
  usevariables = colnames(combined_data),
  rdata = combined_data)

m_step2_fit <- mplusModeler(m_step2,
                 dataout=here("C4-Replication", "Step2_Replication.dat"),
                 modelout=here("C4-Replication", "Step2_Replication.inp") ,
                 check=TRUE, run = TRUE, hashfilename = FALSE)

```

------------------------------------------------------------------------

Step 3

------------------------------------------------------------------------

Estimate multigroup model with thresholds equal across groups
```{r, eval = FALSE}

m_step3  <- mplusObject(
  
  TITLE = "Step 3 - Replication", 
  
  VARIABLE = 
   "categorical = primary-touch; 
    usevar = primary-touch;
  
    knownclass = CG (sample=1 sample=2); !!! LABEL SAMPLES !!!
    classes = CG(2) C(5);                !!! CG is KNOWN CLASS !!!
   ",
  
  ANALYSIS = 
   "estimator = mlr; 
    type = mixture;
    starts = 500 200;
    !STSEED = ; !!! USE SEED TO REPLICATE THESIS RESULTS !!!
   ",
  
  MODEL = 
    "%OVERALL%
     
     !!! INITIAL SAMPLE (SAMPLE 1) !!! INITIAL SAMPLE (SAMPLE 1) !!!
  
     %CG#1.C#1% 

     [ primary$1*-0.21760  ](1);              
     [ change$1*-0.45914   ](2);             
     [ interupt$1*-0.64034 ](3);               
     [ initiat$1*-0.46809  ](4);              
     [ engage$1*-0.33510   ](5);             
     [ approach$1*-0.42163 ](6);               
     [ response$1*-0.58459 ](7);               
     [ expect$1*-0.60368   ](8);             
     [ new$1*1.17714       ](9);         
     [ same$1*2.92160      ](10);         
     [ relative$1*1.86782  ](11);             
     [ objects$1*-1.97843  ](12);             
     [ sequence$1*-1.40431 ](13);              
     [ trans$1*-1.31741    ](14);           
     [ avoid$1*0.99669     ](15);          
     [ control$1*1.69313   ](16);            
     [ touch$1*1.24603     ](17);          

     %CG#1.C#2%

     [ primary$1*3.45112   ](18);        
     [ change$1*2.10148    ](19);       
     [ interupt$1*2.33173  ](20);         
     [ initiat$1*4.67053   ](21);        
     [ engage$1*2.61317    ](22);       
     [ approach$1*3.00689  ](23);         
     [ response$1*3.01867  ](24);         
     [ expect$1*2.47608    ](25);       
     [ new$1*2.13464       ](26);    
     [ same$1*2.61928      ](27);     
     [ relative$1*2.46710  ](28);         
     [ objects$1*3.49010   ](29);        
     [ sequence$1*2.16238  ](30);         
     [ trans$1*1.76331     ](31);      
     [ avoid$1*1.62916     ](32);      
     [ control$1*2.77768   ](33);        
     [ touch$1*2.98188     ](34);      

     %CG#1.C#3%

     [ primary$1*-1.15805  ](35);       
     [ change$1*-1.05640   ](36);      
     [ interupt$1*-1.47052 ](37);        
     [ initiat$1*-1.88494  ](38);       
     [ engage$1*-0.85925   ](39);      
     [ approach$1*-1.29676 ](40);        
     [ response$1*-1.60551 ](41);        
     [ expect$1*-1.84879   ](42);      
     [ new$1*1.21384       ](43);  
     [ same$1*1.25810      ](44);   
     [ relative$1*1.00299  ](45);       
     [ objects$1*2.02420   ](46);      
     [ sequence$1*2.49353  ](47);       
     [ trans$1*1.91476     ](48);    
     [ avoid$1*1.22229     ](49);    
     [ control$1*3.67528   ](50);      
     [ touch$1*2.02070     ](51);    

     %CG#1.C#4%

     [ primary$1*-2.45913  ](52);           
     [ change$1*-2.07089   ](53);          
     [ interupt$1*-2.29483 ](54);            
     [ initiat$1*-2.28241  ](55);           
     [ engage$1*-1.87192   ](56);          
     [ approach$1*-2.25327 ](57);            
     [ response$1*-2.27708 ](58);            
     [ expect$1*-2.29726   ](59);          
     [ new$1*-2.86563      ](60);       
     [ same$1*-2.02751     ](61);        
     [ relative$1*-2.09303 ](62);            
     [ objects$1*-3.25774  ](63);           
     [ sequence$1*-2.18406 ](64);            
     [ trans$1*-2.90318    ](65);         
     [ avoid$1*-2.16937    ](66);         
     [ control$1*-1.95077  ](67);           
     [ touch$1*-1.64226    ](68);         

     %CG#1.C#5%

     [ primary$1*3.92177   ](69);     
     [ change$1*1.18687    ](70);    
     [ interupt$1*2.44270  ](71);      
     [ initiat$1*1.17768   ](72);     
     [ engage$1*4.46855    ](73);    
     [ approach$1*1.51338  ](74);      
     [ response$1*2.11385  ](75);      
     [ expect$1*2.56064    ](76);    
     [ new$1*-2.43669      ](77);  
     [ same$1*-1.04154     ](78);   
     [ relative$1*-1.78324 ](79);       
     [ objects$1*-1.14240  ](80);      
     [ sequence$1*-0.84974 ](81);       
     [ trans$1*-0.89213    ](82);    
     [ avoid$1*-1.11250    ](83);    
     [ control$1*-1.78770  ](84);      
     [ touch$1*-0.87239    ](85); 
  
     !!! REPLICATION SAMPLE (SAMPLE 2) !!! REPLICATION SAMPLE (SAMPLE 2) !!!
  
     %CG#2.C#1% 

     [ primary$1*-0.21760  ](1);             
     [ change$1*-0.45914   ](2);            
     [ interupt$1*-0.64034 ](3);              
     [ initiat$1*-0.46809  ](4);             
     [ engage$1*-0.33510   ](5);            
     [ approach$1*-0.42163 ](6);              
     [ response$1*-0.58459 ](7);              
     [ expect$1*-0.60368   ](8);            
     [ new$1*1.17714       ](9);        
     [ same$1*2.92160      ](10);         
     [ relative$1*1.86782  ](11);             
     [ objects$1*-1.97843  ](12);             
     [ sequence$1*-1.40431 ](13);              
     [ trans$1*-1.31741    ](14);           
     [ avoid$1*0.99669     ](15);          
     [ control$1*1.69313   ](16);            
     [ touch$1*1.24603     ](17);          

     %CG#2.C#2%

     [ primary$1*3.45112   ](18);        
     [ change$1*2.10148    ](19);       
     [ interupt$1*2.33173  ](20);         
     [ initiat$1*4.67053   ](21);        
     [ engage$1*2.61317    ](22);       
     [ approach$1*3.00689  ](23);         
     [ response$1*3.01867  ](24);         
     [ expect$1*2.47608    ](25);       
     [ new$1*2.13464       ](26);    
     [ same$1*2.61928      ](27);     
     [ relative$1*2.46710  ](28);         
     [ objects$1*3.49010   ](29);        
     [ sequence$1*2.16238  ](30);         
     [ trans$1*1.76331     ](31);      
     [ avoid$1*1.62916     ](32);      
     [ control$1*2.77768   ](33);        
     [ touch$1*2.98188     ](34);      

     %CG#2.C#3%

     [ primary$1*-1.15805  ](35);       
     [ change$1*-1.05640   ](36);      
     [ interupt$1*-1.47052 ](37);        
     [ initiat$1*-1.88494  ](38);       
     [ engage$1*-0.85925   ](39);      
     [ approach$1*-1.29676 ](40);        
     [ response$1*-1.60551 ](41);        
     [ expect$1*-1.84879   ](42);      
     [ new$1*1.21384       ](43);  
     [ same$1*1.25810      ](44);   
     [ relative$1*1.00299  ](45);       
     [ objects$1*2.02420   ](46);      
     [ sequence$1*2.49353  ](47);       
     [ trans$1*1.91476     ](48);    
     [ avoid$1*1.22229     ](49);    
     [ control$1*3.67528   ](50);      
     [ touch$1*2.02070     ](51);    

     %CG#2.C#4%

     [ primary$1*-2.45913  ](52);       
     [ change$1*-2.07089   ](53);      
     [ interupt$1*-2.29483 ](54);        
     [ initiat$1*-2.28241  ](55);       
     [ engage$1*-1.87192   ](56);      
     [ approach$1*-2.25327 ](57);        
     [ response$1*-2.27708 ](58);        
     [ expect$1*-2.29726   ](59);      
     [ new$1*-2.86563      ](60);   
     [ same$1*-2.02751     ](61);    
     [ relative$1*-2.09303 ](62);        
     [ objects$1*-3.25774  ](63);       
     [ sequence$1*-2.18406 ](64);        
     [ trans$1*-2.90318    ](65);     
     [ avoid$1*-2.16937    ](66);     
     [ control$1*-1.95077  ](67);       
     [ touch$1*-1.64226    ](68);     

     %CG#2.C#5%

     [ primary$1*3.92177   ](69);     
     [ change$1*1.18687    ](70);    
     [ interupt$1*2.44270  ](71);      
     [ initiat$1*1.17768   ](72);     
     [ engage$1*4.46855    ](73);    
     [ approach$1*1.51338  ](74);      
     [ response$1*2.11385  ](75);      
     [ expect$1*2.56064    ](76);    
     [ new$1*-2.43669      ](77);  
     [ same$1*-1.04154     ](78);   
     [ relative$1*-1.78324 ](79);       
     [ objects$1*-1.14240  ](80);      
     [ sequence$1*-0.84974 ](81);       
     [ trans$1*-0.89213    ](82);    
     [ avoid$1*-1.11250    ](83);    
     [ control$1*-1.78770  ](84);      
     [ touch$1*-0.87239    ](85);   ",
  
  
  OUTPUT = "",

  PLOT = "type = plot3; 
          series = primary-touch(*);",
  
  usevariables = colnames(combined_data),
  rdata = combined_data)

m_step3_fit <- mplusModeler(m_step3,
                 dataout=here("C4-Replication", "Step3_Replication.dat"),
                 modelout=here("C4-Replication", "Step3_Replication.inp") ,
                 check=TRUE, run = TRUE, hashfilename = FALSE)

```

------------------------------------------------------------------------

Step 4: Compare model fit

------------------------------------------------------------------------

- Unconstrained model: `Step2_Replication.out` (no threshold constraints)

- Constrained model: `Step3_Replication.out` (thresholds constrained to equality across groups)

```{r}
replication_models <- readModels(here("C4-Replication"), quiet = TRUE)
```

Compare model fit summary statistics: Exploratory & Confirmatory Models 
```{r, eval=TRUE}

enum_extract1 <- LatexSummaryTable(replication_models$Step2_Replication.out ,                                 
                 keepCols=c("Title", "Parameters", "LL", "BIC", "aBIC", "Observations")) 
enum_extract2 <- LatexSummaryTable(replication_models$Step3_Replication.out,                                 
                 keepCols=c("Title", "Parameters", "LL", "BIC", "aBIC", "Observations")) 
```

Calculate indices derived from the Log Likelihood (LL)
```{r, eval=TRUE}
                           
allFit <- rbind(enum_extract1, enum_extract2) %>% 
  mutate(aBIC = -2*LL+Parameters*log((Observations+2)/24)) %>% 
  select(1:5) %>% 
  mutate(Title = case_when(
    Title == " Step 2 - Replication" ~ "Unconstrained model",
    Title == " Step 3 - Replication" ~ "Constrained model"))

```

Format fit table
```{r, eval=TRUE}

allFit %>% 
  gt() %>%
  tab_header(
    title = md("**Model Fit Comparision Table**"), subtitle = md("&nbsp;")) %>% 
  cols_label(
    Title = "Model",
    Parameters = md("Par"),
    LL = md("*LL*"),
    BIC = md("BIC"),
    aBIC = md("aBIC")) %>% 
  tab_footnote(
    footnote = md(
    "*Note.* Par = Parameters; *LL* = model log likelihood;
     BIC = Bayesian information criterion; aBIC = sample size adjusted BIC."), 
    locations = cells_title()) %>% 
  tab_options(column_labels.font.weight = "bold")
```








